---
title: "RStudio ETL, EDA, and Control Tests Notebook"
output:
  html_document:
    df_print: paged
---

************************************************************************
Title: "RStudio ETL, EDA, and Control Tests Notebook"
************************************************************************

# Load packages
```{r}
library(tidyverse)
library(lubridate)
```

# ETL - Extract, Transform, and Load the data
## Load data
```{r}
dfa <- read_csv("pcard_data_part1_v01.csv")

dfb <- read_csv("pcard_data_part1_v01.csv")
```

## Join the data together before fixing the column headings
```{r}
dftemp <- bind_rows(dfa, dfb)
```

## Rename the `MCC.Description` column in dfa and join correctly
```{r}
# Rename
dfa <- dfa %>% rename(MCC = MCC.Description)

# join
df <- bind_rows(dfa, dfb)
```

## Filter to keep only OSU
```{r}
df <- df %>% filter(Agency.Number == 1000)
```

## Modify all character columns to make the strings all uppercase
```{r}
df <- df %>% mutate(across(where(is.character), str_to_upper))
```

## 1) Eliminate "Project"; 2) Convert columns that should be numbers; 3) Convert columns that should be dates
```{r}
# Eliminate `Project`
##df <- df %>% select(-Project)

# numeric and date columns are already correct
str(df)
```

## Add three date columns
```{r}
df <- df %>% mutate(Month = month(Transaction.Date))
df <- df %>% mutate(Weekday = wday(Transaction.Date))
df <- df %>% mutate(Day.Month = mday(Transaction.Date))
```

# Exploratory Data Analysis
## Distribution statistics for `Amount`
```{r}
summary(df$Amount)
```

## Box plot for `Amount`
```{r}
df %>% ggplot(aes(y=Amount)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1000, 1000))
```

## Sum of `Amount` by month
```{r}
des2 <- df %>% group_by(Month) %>%  
  summarize(sum_amount = sum(Amount))

des2 %>% ggplot(aes(y=sum_amount, x=as.factor(Month), fill=as.factor(Month))) + 
  geom_col()
```

## Sum of `Amount` by day of the week
```{r}
df %>% 
  group_by(Weekday) %>%  
  summarize(sum_amount = sum(Amount)) %>% 
  ggplot(aes(y=sum_amount, x=as.factor(Weekday), fill=as.factor(Weekday))) + 
  geom_col()
```

## Sum of `Amount` by day of the month
```{r}
df %>% group_by(Day.Month) %>%  summarize(sum_amount = sum(Amount)) %>% 
  ggplot(aes(y=sum_amount, x=as.factor(Day.Month), fill=as.factor(Day.Month))) + 
  geom_col()
```

# Tests of controls
## Single Transaction Limit, descending by `Amount`
```{r}
## All transactions
df %>% filter(Amount>5000) %>% 
  arrange(desc(Amount))
```

## Single transaction limit, ascending by `Cardholder.Name`
```{r}
## All transactions
df %>% filter(Amount>5000) %>% 
  arrange(Cardholder.Name)
```

## Single transaction limit, ascending by `Item.Description`
```{r}
## All transactions
df %>% filter(Amount>5000) %>% 
  arrange(Item.Description)
```

## Single transaction limit, by person
```{r}
df %>% filter(Amount>5000) %>% 
  group_by(Cardholder.Name) %>% 
  summarize(count = n(), sum_amount=sum(Amount), avg_amount=mean(Amount)) %>% 
  arrange(desc(count))
```

## Yearly spending limit
```{r}
df %>% group_by(Cardholder.Name) %>% 
  summarize(sum_amount = sum(Amount), count = n()) %>% 
  filter(sum_amount >50000) %>% 
  arrange(desc(sum_amount))
```

## Monthly spending limit violations by person
```{r}
df %>% group_by(Cardholder.Name, Month) %>% 
  summarize(sum_amount = sum(Amount)) %>% 
  filter(sum_amount >10000) %>% 
  summarize(count_months_over = n(), total_spent = sum(sum_amount), average_per_month=mean(sum_amount)) %>% 
  arrange(desc(count_months_over))
```
